---
title: "A2-MOTAENGIL"
author: "helder"
date: "`r Sys.Date()`"
output: pdf_document
---

## MOTAENGIL {-}

In figure \@ref(fig:nvb-acf-pacf), the ACF plots show no significant correlations. ARCH-LM test in table \@ref(tab:nvb-arch-test) does support evidence of ARCH effect (the null hypothesis of no existence of ARCH effect is rejected), and also the ACF of the squared series have a significant correlation at lag 1. The Ljung-Box test, in table \@ref(tab:nvb-ljung-box-test) is rejected for significant autocorrelations until lag 12.

The candidate models goodness of fit (gof) summary table \@ref(tab:nvb-fits), will lead to select the ARMA(1,1)-GARCH(1,1)-M type of model. Even thought, ACF plots did not support the existance of an ARMA process for the mean, the auto selection predicted it, and the following results will conclude the significance of the coeficients. Keeping the rational of selecting the model based on best AIC.

The model coefficients are resumed in the table \@ref(tab:nvb-model-coef-tab). The mu and omega are estimated zero. All the other coefficients are significant. The archm coefficient is negative indicating that the negatively related to its past volatility.

The standardized residuals plots in figure \@ref(fig:nvb-residual-plots) confirm the appropriate choice of t-distribution. Other diagnostic test checks are confirmed, the Ljung–Box and The LM Arch tests, all lead to conclude no correlation and no ARCH effect in the residuals. The Jarque-Bera test also conclude the residuals are normally distributed.

The model APARCH(1, 1) with t-distribution was used to forecast the volatility ploted in figure \@ref(fig:nvb-forecast-plots).

```{r}
NAME <- "MOTAENGIL"
s <- log_returnMOTAENGIL
```

```{r mot-acf-pacf, fig.pos="!h", out.extra = "", fig.cap=paste(NAME, "Log Returns"), fig.dim=c(6, 4), include=TRUE}
p1 <- as_tsibble(s) %>% ACF() %>% autoplot() + ylim(c(-0.10, 1)) + ggtitle("log returns")
p2 <- as_tsibble(s) %>% PACF() %>% autoplot() + ylim(c(-0.10, 1))
p3 <- as_tsibble(s^2) %>% ACF() %>% autoplot() + ylim(c(-0.10, 1)) + ggtitle("log returns ^2")
p4 <- as_tsibble(s^2) %>% PACF() %>% autoplot() + ylim(c(-0.10, 1))
(p1 + p2) / (p3 + p4)
```

```{r mot-arch-test}
kbl(do_arch_test(s),
      caption = paste(NAME, "- ARCH LM test"),
      booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r mot-ljung-box-test}
kbl(tidy(Box.test(coredata(s^2), type="Ljung-Box", lag = 12)),
      caption = paste(NAME, "- Ljung-Box test"),
      booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r mot-fits-00, include=TRUE}
arma.order <- c(0,0)

arch.order <-  1:2
arch.names <-  paste("arch", arch.order, sep="")
fit.list <-  list()
# for (p in arch.order) {
#   arch.spec <-  ugarchspec(variance.model = list(garchOrder=c(p,0)),
#                           mean.model = list(armaOrder=arma.order))
#   fit.list[[p]] <- ugarchfit(spec=arch.spec, data=s)
# }
# names(fit.list) <-  arch.names

garch11.spec <-  ugarchspec(variance.model=list(garchOrder=c(1,1)),
                          mean.model = list(armaOrder=arma.order))
fit.list$garch11 <- ugarchfit(spec=garch11.spec, data=s)

garch11_std.spec <-  ugarchspec(variance.model=list(garchOrder=c(1,1)),
                          mean.model = list(armaOrder=arma.order), distribution.model="std")
fit.list$garch11_std <- ugarchfit(spec=garch11_std.spec, data=s)

garch22_std.spec <-  ugarchspec(variance.model=list(garchOrder=c(2,2)),
                              mean.model = list(armaOrder=arma.order), distribution.model="std")

fit.list$garch22_std <- ugarchfit(spec=garch22_std.spec, data=s)

igarch11_std.spec <-  ugarchspec(variance.model=list(model="iGARCH", garchOrder=c(1,1)),
                             mean.model = list(armaOrder=arma.order, include.mean=TRUE),
                             distribution.model="std")

fit.list$igarch11_std <- ugarchfit(spec=igarch11_std.spec, data=s)

garch_m_11_std.spec  <-  ugarchspec(variance.model = list(model="sGARCH", garchOrder=c(1,1)),
                                mean.model = list(armaOrder=arma.order, 
                                                  include.mean=TRUE, archm=TRUE, archpow = 2),
                                distribution.model="std")
fit.list$garch_m_11_std <- ugarchfit(spec=garch_m_11_std.spec, data=s)

apgarch11_std.spec = ugarchspec(variance.model = list(model="apARCH", garchOrder=c(1,1)),
                                mean.model = list(armaOrder=arma.order, include.mean=TRUE),
                                distribution.model="std")

aparch11_std.spec = ugarchspec(variance.model = list(model="apARCH", garchOrder=c(1,1)),
                                mean.model = list(armaOrder=arma.order, include.mean=TRUE),
                                distribution.model="std")

fit.list$aparch11_std <- ugarchfit(spec=aparch11_std.spec, data=s)
info.mat <- sapply(fit.list, infocriteria)
rownames(info.mat) <-  rownames(infocriteria(fit.list[[1]]))

kbl(info.mat[1:2,],
      caption = paste(NAME, " ARMA(0,0)-*ARCH(p,q) - Models goodness of fit (AIC; BIC)"),
      booktabs = T) %>%
  kable_styling(latex_options = c("hold_position","scale_down"))
```

```{r mot-fits, include=TRUE}
arma.order <- c(1,1)

arch.order <-  1:2
arch.names <-  paste("arch", arch.order, sep="")
fit.list <-  list()
# for (p in arch.order) {
#   arch.spec <-  ugarchspec(variance.model = list(garchOrder=c(p,0)),
#                           mean.model = list(armaOrder=arma.order))
#   fit.list[[p]] <- ugarchfit(spec=arch.spec, data=s)
# }
# names(fit.list) <-  arch.names

garch11.spec <-  ugarchspec(variance.model=list(garchOrder=c(1,1)),
                          mean.model = list(armaOrder=arma.order))
fit.list$garch11 <- ugarchfit(spec=garch11.spec, data=s)

garch11_std.spec <-  ugarchspec(variance.model=list(garchOrder=c(1,1)),
                          mean.model = list(armaOrder=arma.order), distribution.model="std")
fit.list$garch11_std <- ugarchfit(spec=garch11_std.spec, data=s)

garch22_std.spec <-  ugarchspec(variance.model=list(garchOrder=c(2,2)),
                              mean.model = list(armaOrder=arma.order), distribution.model="std")

fit.list$garch22_std <- ugarchfit(spec=garch22_std.spec, data=s)

igarch11_std.spec <-  ugarchspec(variance.model=list(model="iGARCH", garchOrder=c(1,1)),
                             mean.model = list(armaOrder=arma.order, include.mean=TRUE),
                             distribution.model="std")

fit.list$igarch11_std <- ugarchfit(spec=igarch11_std.spec, data=s)

garch_m_11_std.spec  <-  ugarchspec(variance.model = list(model="sGARCH", garchOrder=c(1,1)),
                                mean.model = list(armaOrder=arma.order, 
                                                  include.mean=TRUE, archm=TRUE, archpow = 2),
                                distribution.model="std")
fit.list$garch_m_11_std <- ugarchfit(spec=garch_m_11_std.spec, data=s)

aparch11_std.spec = ugarchspec(variance.model = list(model="apARCH", garchOrder=c(1,1)),
                                mean.model = list(armaOrder=arma.order, include.mean=TRUE),
                                distribution.model="std")

fit.list$aparch11_std <- ugarchfit(spec=aparch11_std.spec, data=s)

info.mat <- sapply(fit.list, infocriteria)
rownames(info.mat) <-  rownames(infocriteria(fit.list[[1]]))

kbl(info.mat[1:2,],
      caption = paste(NAME, " ARMA(1,1)-*ARCH(p,q) - Models goodness of fit (AIC; BIC)"),
      booktabs = T) %>%
  kable_styling(latex_options = c("hold_position","scale_down"))
```


```{r mot-model-coef-tab}
#arch.list$arch1@fit$matcoef
#arch.list$garch11@fit$matcoef
kbl(round(fit.list$garch_m_11_std@fit$matcoef, 4),
      caption = paste(NAME, "GARCH-M - Model coefficients"),
      booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r, eval=FALSE}
show(fit.list$garch_m_11_std)
```

Model diagnostics:
\singlespacing
\begin{verbatim}
GARCH Model	: sGARCH(1,1)   Mean Model	: ARFIMA(1,0,1)   Distribution	: std 
Weighted Ljung-Box Test on Standardized Residuals
                        statistic p-value
Lag[1]                     0.1211  0.7278
Lag[2*(p+q)+(p+q)-1][5]    0.2509  1.0000
Lag[4*(p+q)+(p+q)-1][9]    0.8374  0.9999
d.o.f=2       H0 : No serial correlation
Weighted Ljung-Box Test on Standardized Squared Residuals
                        statistic p-value
Lag[1]                      2.894 0.08891
Lag[2*(p+q)+(p+q)-1][5]     3.869 0.27084
Lag[4*(p+q)+(p+q)-1][9]     4.431 0.51746
d.o.f=2
Weighted ARCH LM Tests
            Statistic Shape Scale P-Value
ARCH Lag[3]     1.090 0.500 2.000  0.2964
ARCH Lag[5]     1.332 1.440 1.667  0.6375
ARCH Lag[7]     1.605 2.315 1.543  0.8000
```{r , results='asis'}
jb <- rugarch::BerkowitzTest(residuals(fit.list$garch_m_11_std, standardize=TRUE))[10:11]
print(paste("Jarque-Bera Test of Normality Statistic:", round(jb$JB, 4), " p-value =", round(jb$JBp, 4) ))
```
\end{verbatim}
\onehalfspacing

```{r mot-residual-plots, fig.pos="!h", out.extra = "", fig.cap=paste(NAME, "GARCH-M Model Standardized Residuals"), fig.dim=c(6, 3), include=TRUE}
par(mfrow = c(1, 2))
plot(fit.list$garch_m_11_std, which=8) #8:   Empirical Density of Standardized Residuals
plot(fit.list$garch_m_11_std, which=9) #9:   QQ-Plot of Standardized Residuals
```


```{r mot-forecast-plots, fig.pos="!h", out.extra = "", fig.cap=paste(NAME, "GARCH-M Model Forecasts"), fig.dim=c(6, 3), include=TRUE}
par(mfrow = c(1, 2))
garchforecast1 <- ugarchforecast(fit.list$garch_m_11_std, data=s, n.ahead = 5)
plot(garchforecast1, which=1) #1:   Time Series Prediction (unconditional)
plot(garchforecast1, which=3) #3:   Sigma Prediction (unconditional)
```

