---
title: "Antas-Espinho"
author: "helder"
date: "`r Sys.Date()`"
output: pdf_document
---

## Antas-Espinho {-}

```{r , include=FALSE}
y <- qualityAR03 %>% filter(Location=="Antas-Espinho")
```

The Antas-Espinho series is ploted in the figure \@ref(fig:ae-ts). By visual inspection of the series, and the ACF plot is clear that there is a strong seasonal component at lags 24, 48, 72, etc, and the series is not stationary in season (ACF decays slowly). An Order 1 Difference in seasonal component of SARIMA model will be used. The are some high values, far from the usual variance band of the series, so to improve stability and homosedasticity a Box Cox transformation will be applied.

Applying Unit root tests as resumed in table \ref{tab:ae-test-stationary}, one may be inclined to follow a differencing own frequency instead seasonal differencing. The two tests have opposing decisions, it can follow from the fact that the series is very long and the use of high number of points can sometimes lead to null hypothesis rejection. It is clear by ACF and PACF inspection that there is need to do seasonal difference.

```{r ae-ts, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Time series"), fig.dim=c(7, 3), include=TRUE}
y %>% gg_tsdisplay(Ozone, plot_type = "partial", lag_max = 200)
```

```{r, echo=FALSE, include=FALSE}
lambda <- y %>%
  features(Ozone, features = guerrero) %>%
  pull(lambda_guerrero)
#caption_tex = paste0("Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2))
```

```{r ae-test-stationary, include=TRUE}
adf <- function(x, ...) {
  out <- tseries::adf.test(x, k = 1)
  c(adf_stat = unname(out$statistic), adf_pvalue = out$p.value)
}
# tseries::adf.test(box_cox(y$Ozone, lambda), k=1)
kbl(y %>%
      mutate(Ozone = box_cox(Ozone, lambda)) %>%
      features(Ozone, 
               list(unitroot_kpss, adf, unitroot_ndiffs, ~ unitroot_nsdiffs(., .period = 24))),
    caption = paste(y$Location[1], " - Unit root tests and suggested number of differences"),
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Looking at the seasonal difference series in figure \@ref(fig:ae-diffplot), it shows no trend, and the ACF and PACF plots are typical of a stationary series, and there is no need to further differentiate. There is only one seasonal significant correlation at lag 24 in the ACF meaning that an SMA(1) component model will be used, also only the 3 first lags in PACF are significant (maybe lag 3 can be considered non-significant) and a AR(2) component will be considered in the final predicted SARIMA model (2,0,0)(0,1,1).

```{r ae-diffplot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2)), fig.dim=c(7, 4), include=TRUE}
y %>%
  gg_tsdisplay(difference(box_cox(Ozone, lambda), lag = 24), plot_type = "partial", lag_max = 200) +
  labs(y = "")
```

```{r , include=FALSE}
fits <- readRDS("models/antas-espinho_fit_s011.rds")
fit <- fits %>% select(Location, sarima101011)
```

Diagnostics in figure \@ref(fig:ae-fit-resid-plot) show that the model does not pass the test for non-correlated residuals.

```{r, include=FALSE}
plot.box.ljung <- function (
    z, k = 10, main = "p-value of the Ljung-Box test", ylab = "p-value"
) {
  p <- rep(NA, k)

  for (i in 1:k) {
    p[i] <- Box.test(z, i, type = "Ljung-Box")$p.value
  }

  ggplot(as_tibble(p), aes(x=1:k, y=p)) +
    geom_bar(stat='identity', width =0.05) +
    geom_point() + ylim(0,1) + ylab(ylab) +
    geom_abline(intercept = .05, slope = 0, lty = 3) +
    ggtitle(main)
}

fit_diag_plots <- function(fit) {
  m <- fit %>% augment() %>% mutate( .innov = scale(.innov) )

  p1 <- m %>% autoplot(.innov)
  p2 <- m %>% ACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p3 <- m %>% PACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p4 <- m %>%  gghistogram(x=".innov", bins = 150)
  p5 <- plot.box.ljung(m %>% pull(.innov), k=10)
  p6 <- ggqqplot(m,".innov", title = "QQ plot")
  p7 <- ggplot(m, aes(x = .fitted, y = .innov)) +
          geom_point()
  
  (p1) / (p2 + p3) / (p4 + p5) / (p6 + p7)
}

```


```{r ae-fit-resid-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE}
fit_diag_plots(fit)
```

Variations of predicted SARIMA model where estimated, resumed in table  \@ref(tab:ae-model-comparison), to evaluate if there are improvements in model diagnostics requirement fulfillment. The best overall model is the `r format(fit %>% pull(2))`. (the sarima100011 was also tried but failed the uncorrelation tests of the residuals (ljung-box)).

Proceeding to model diagnostics, in \@ref(fig:ae-fit-resid-plot) is shown that the residuals are uncorrelated, Ljung-box tests all reject null hypotesys of correlation until lag 10, but the QQ plot show that there's a divergence from normality in the tails of the distribution. The Fitted vs Residual plot does not show any pattern of concern (having greater errors for greater fitted values for example).

```{r ae-model-comparison, include=TRUE}
kbl(fits %>% 
      pivot_longer(!Location, names_to = "Model name",
                     values_to = "Orders") %>%
      mutate("Model" = format(Orders)) %>%
      glance() %>%
      select(Location, Model, AIC, AICc, BIC),
    format = "latex", 
    caption = paste(y$Location[1], "model comparison"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

The model `r format(fit %>% pull(2))` coefficients are statistical significant, as resumed in table \@ref(tab:ae-coef), all p-values of the t-test (H0: coefficient is zero) lead to rejecting the null hypothesis. The process evolves with positive correlation with the last value and negative correlation with last seasonal period.

```{r ae-coef, include=TRUE}
kbl(fit %>% coef(), 
    format = "latex", 
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Coefficient statistics"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Model forecasts are resumed in the table \@ref(tab:ae-forecast-tab), and figure \@ref(fig:ae-forecast-plot). The confidence interval of the prediction gets wither with for each forecasted new period, representing the added uncertainty of the forecasts.

Cross-validation on the last month shows the model is not capturing the variance amplitude of the series, true values have more dynamic range than the forecast (horizon=5), this is expected because of increasing the horizon leads to the forecast being closer to the expected value of the distribution. Also the residuals of the model do not have strong resemblance to normal distribution, therefore it is expected more extreme values in the real data than those predicted by the model.

```{r ae-forecast-tab, include=TRUE}
f <- fit %>% forecast(h=5)

knitr::kable(f %>%  hilo(level = c(95)) %>% select(-c(1,2)),
             caption = paste(y$Location[1], format(fit %>% pull(2)), "forecast 5 periods ahead (95\\% CI)"),
             booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r ae-forecast-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "forecasts"), fig.dim=c(6, 2), include=TRUE}
f %>% 
  autoplot(y %>%
              filter(between(index,as_datetime("2020-12-15"), as_datetime("2021-01-01"))),
            level = 95,
            alpha = 0.5)
```

```{r ae-cv, include=FALSE}
training <- y %>% filter(month(index) <= 11)
test <- y %>% filter(month(index) > 11)

train_fit <- training %>%
  model(
    sarimacv = ARIMA(box_cox(Ozone, lambda) ~ 0 + pdq(1, 0, 1) + PDQ(0, 1, 1, period=24))
  )

test_refit <- train_fit %>% refit(test)
# fits5 <- test_refit %>% fitted(h = 5)
# saveRDS(fits5, file = "models/antas-espinho_cv_fits5.rds")
# Load
fits5 <- readRDS("models/antas-espinho_cv_fits5.rds")

#fits1 <- test_refit %>% fitted(h = 1)
```

```{r ae-cv-accuracy-tab}
kbl(bind_rows(
    fit %>% accuracy() %>% select(RMSE, MAE),
    test_refit %>% accuracy() %>% select(RMSE, MAE)
    ) %>% 
      add_column(Type = c("Series", "CV")) %>% 
      select(3, 1, 2),
    caption = paste(y$Location[1], format(test_refit %>% pull(2)), "Series and Crossvalidation Accuracy"),
             booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r ae-cv-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(test_refit %>% pull(2)), "crossvalidation forecasts"), fig.dim=c(6, 2), include=TRUE}
test %>%
  autoplot(Ozone) +
  autolayer(fits5, .fitted, col = "#D55E00")
```

