---
title: "SARIMA models of ground level Ozone concentrations in 10 Portuguese locations"
author: "Helder Jose Sousa Dias #46541"
course: "Course: Time Series 2021/2022"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_book:
    template: projectreport.tex
    toc: no
    number_sections: no
    keep_tex: yes
    fig_caption: yes
bibliography: references.bib
link-citations: true
biblio-style: unsrt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	include = FALSE,
	cache = TRUE,
	eval.after = "fig.cap"
	#fig.pos = "!H"
	#fig.env="figure"
)

options(knitr.table.format = "latex") # For kable tables to write LaTeX table directly
```

# Abstract {-}

Air quality is a major factor in life expectancy. Particularly in urban areas government authorities need to control air quality and take measures if human health is in jeopardy. Ozone concentrations at ground level are one of air quality measures. The SARIMA models were designed to model not only time dependency structure of series along past values but also encompass seasonal components, characteristic of recurrent patterns in hourly, daily, etc periods. By fitting a SARIMA model to the series short term forecasts can be obtained and used to predict the probability of extreme events. The 10 series studied showed daily seasonality and autoregressive component, confirming accordance to known fact of concentrations fluctuation along the day and being physical quantities the present if directly correlated with previous value. It was not devised a model form equal to all locations, but some similarities were identified, in Porto region leading to a SARIMA(1,0,1)(0,1,1)[24] form as the most common among the locations.

```{r imports}
library(tidyverse, quietly = TRUE)
library(fpp3, quietly = TRUE)
library(imputeTS)
library(latex2exp)
library(patchwork)
library(ggpubr)
library(kableExtra)
```

```{r loadData}
qualityAR03 <- readxl::read_excel("./data/QualidadeARO3.xlsx") %>% 
  ts(.,start=c(2020,1,1),frequency=24*366) %>% 
  as_tsibble() %>% 
  rename(Location = key, Ozone = value) %>% 
  mutate(index = as_datetime(round_date(index, unit = "hour")))
```

```{r child = 'A1-Introduction.Rmd', include=TRUE}
```

```{r child = 'A1-ModelingWorkflow.Rmd', include=TRUE}
```

# Results & Discussion {-}
\label{sec:results}

```{r child = 'Antas-Espinho.Rmd', include=TRUE}
```

```{r child = 'Entrecampos.Rmd', include=TRUE}
```

```{r child = 'Estarreja.Rmd', include=TRUE}
```

```{r child = 'Ilhavo.Rmd', include=TRUE}
```

```{r child = 'Laranjeiro-Almada.Rmd', include=TRUE}
```

```{r child = 'Mem-Martins.Rmd', include=TRUE}
```

```{r child = 'Paio-Pires.Rmd', include=TRUE}
```

```{r child = 'Restelo.Rmd', include=TRUE}
```

```{r child = 'Sobreiras-Porto.Rmd', include=TRUE}
```

```{r child = 'VNTelha-Maia.Rmd', include=TRUE}
```

# Conclusions {-}

Modeling workflow was followed to select the best SARIMA model for each of the 10 locations, this was a time daunting task, and this report does not make justice.

ACF and PACF analysis proved to be good tools to select model orders. Model diagnostics is a very important step that was difficult to overcome for all the series. All the models estimated failed to produce normally distributed residuals. The format of the series may justify this limitation. Obtaining non-correlated residuals was also difficult in most cases, and the predicted model by ACF / PACF inspection failed in often in this requirement.

In this type of model, with differences applied, was not possible to interpret the coefficients to have some intuitive understanding of the natural process.

It was not possible to select one model that fits all the locations, but some locations share the same form namely Antas-Espinho, Sobreiras-Porto, and VNTelha-Maia all in the Porto region (within 20 Km distance).

Left for future work, due to time constraints, were ideas of using the models for computing the probability of extreme events, and also study timeseries clustering for evaluating similarities in the locations.

# Bibliography {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\noindent
<div id="refs"></div>
