---
title: "Entrecampos"
author: "helder"
date: "`r Sys.Date()`"
output: pdf_document
---

## Entrecampos {-}

```{r , include=FALSE}
y <- qualityAR03 %>% filter(Location=="Entrecampos")
```

```{r, echo=FALSE, include=FALSE}
lambda <- y %>%
  features(Ozone, features = guerrero) %>%
  pull(lambda_guerrero)
#caption_tex = paste0("Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2))
```

The Entrecampos series is ploted in figure \@ref(fig:entr-ts). The ACF an PACF plots show non-stationarity, and a seasonal component at lags 24, 48, etc. To test if should it be applied regular difference or a seasonal difference, was tested a regular difference, but the ACF plot still showed non-stationary pattern. An Order 1 Difference in seasonal component of SARIMA model will be used. The are fluctuations in the variance, so to improve stability and homoscedasticity a Box Cox transformation will be applied.

Again, applying Unit root tests as resumed in table \ref{tab:entr-test-stationary} leads to incorrect assumptions, of for what type of difference should by applied.

```{r entr-ts, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Time series"), fig.dim=c(7, 3), include=TRUE}
y %>% gg_tsdisplay(Ozone, plot_type = "partial", lag_max = 200)
```

```{r entr-test-stationary, include=TRUE}
adf <- function(x, ...) {
  out <- tseries::adf.test(x, k = 1)
  c(adf_stat = unname(out$statistic), adf_pvalue = out$p.value)
}
# tseries::adf.test(box_cox(y$Ozone, lambda), k=1)
kbl(y %>%
      mutate(Ozone = box_cox(Ozone, lambda)) %>%
      features(Ozone, 
               list(unitroot_kpss, adf, unitroot_ndiffs, ~ unitroot_nsdiffs(., .period = 24))),
    caption = paste(y$Location[1], " - Unit root tests and suggested number of differences"),
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Looking at the seasonal difference series (h=24) in figure \@ref(fig:entr-diffplot), it shows no trend, and the ACF and PACF plots look typical of a stationary series (no further differentiate). There are two seasonal significant correlation at lag 24 and 48 in the ACF, meaning that an SMA(2) component model will be used. In the the first lags only 2 are significant in the PACF and a AR(2) component will be considered. The final predicted model is of the form SARIMA(2,0,0)(0,1,2)[24]. Further analysis of this model, lead to conclude the residuals have some correlation at higher lags, therefore variations around this form were considered.
Goodness-of-fit model comparison is resumed in table \@ref(tab:entr-model-comparison).

```{r entr-diffplot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2)), fig.dim=c(7, 4), include=TRUE}
y %>%
  gg_tsdisplay(difference(box_cox(Ozone, lambda), lag = 24), plot_type = "partial", lag_max = 200) +
  labs(y = "")
```

```{r , include=FALSE}
#  [1] "Location"     "sarima000101" "sarima100101" "sarima001101" "sarima101101" "sarima201101"
#  [7] "sarima102101" "sarima202101" "sarima000012" "sarima100012" "sarima200012" "sarima101012"
# [13] "sarima201012"
#fits <- readRDS("models/entrecampos_fit_s012.rds") %>% select(-c(2, 3, 4, 5, 6, 7, 8, 9))
fits <- readRDS("models/entrecampos_fit_s011.rds")
fit <- fits %>% select(Location, sarima101012)
```

Diagnostics in figure \@ref(fig:entr-fit-resid-plot) show that the model does not clearly pass the test for non-correlated residuals, for higher lags. Again the QQ plot shows patterns diverging from normality in the tails of the distribution. The Fitted vs Residual plot does not show any pattern of concern.

The alternative <ARIMA(1,0,1)(0,1,1)[24]> does not lead to non correlated residuals. The decision was to keep this model `r format(fit %>% pull(2))`, since the same pattern is present in other alternatives, and the inspection of ACF and PACF did not show any signs of other significant correlations in the series.

Also an alternative difference form (d=1, D=1), to check if the autocorrelation in lag 48 was not significant but resulted in more strong residual correlations.

```{r, include=FALSE}
plot.box.ljung <- function (
    z, k = 10, main = "p-value of the Ljung-Box test", ylab = "p-value"
) {
  p <- rep(NA, k)

  for (i in 1:k) {
    p[i] <- Box.test(z, i, type = "Ljung-Box")$p.value
  }

  ggplot(as_tibble(p), aes(x=1:k, y=p)) +
    geom_bar(stat='identity', width =0.05) +
    geom_point() + ylim(0,1) + ylab(ylab) +
    geom_abline(intercept = .05, slope = 0, lty = 3) +
    ggtitle(main)
}


fit_diag_plots <- function(fit) {
  m <- fit %>% augment() %>% mutate( .innov = scale(.innov) )

  p1 <- m %>% autoplot(.innov)
  p2 <- m %>% ACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p3 <- m %>% PACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p4 <- m %>%  gghistogram(x=".innov", bins = 150)
  p5 <- plot.box.ljung(m %>% pull(.innov), k=10)
  p6 <- ggqqplot(m,".innov", title = "QQ plot")
  p7 <- ggplot(m, aes(x = .fitted, y = .innov)) +
          geom_point()
  
  (p1) / (p2 + p3) / (p4 + p5) / (p6 + p7)
}

```

```{r entr-fit-resid-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE}
fit_diag_plots(fit)
```

```{r entr-model-comparison, include=TRUE}
kbl(fits %>% 
      pivot_longer(!Location, names_to = "Model name",
                     values_to = "Orders") %>%
      mutate("Model" = format(Orders)) %>%
      glance() %>%
      select(Location, Model, AIC, AICc, BIC),
    format = "latex", 
    caption = paste(y$Location[1], "model comparison"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

```{r}
fit <- fits %>% select(Location, sarima101012)
```

The coefficients are statistical significant, as resumed in table \@ref(tab:entr-coef), all p-values of the t-test (H0: coefficient is zero) lead to rejecting the null hypothesis. Also in this location the process evolves with positive correlation with the last value and negative correlation with last seasonal period.

```{r entr-coef, include=TRUE}
kbl(fit %>% coef(), 
    format = "latex", 
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Coefficient statistics"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Model forecasts were obtained in the table \@ref(tab:entr-forecast-tab), and figure \@ref(fig:entr-forecast-plot). The forecast plot (that show only a small part of the series) shows that there are several values close to zero and consecutive periods, this is a trait that can justify the abnormal pattern in the residuals of the model, because no normal stochastic process can generate this kind of pattern.

To test the model when applying cross-validation (with horizon=5) the model seems to follows the true path, having some discrepancies, but maintaining the dynamic range of the series. In a real world setup the forecasts should be truncated of values bellow zero.

```{r entr-forecast-tab, include=TRUE}
f <- fit %>% forecast(h=5)

kbl(f %>%  hilo(level = c(95)) %>% select(-c(1,2)),
             caption = paste(y$Location[1], format(fit %>% pull(2)), "forecast 5 periods ahead (95\\% CI)"),
             booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r entr-forecast-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "forecasts"), fig.dim=c(6, 2), include=TRUE}
f %>% 
  autoplot(y %>%
              filter(between(index,as_datetime("2020-12-15"), as_datetime("2021-01-01"))),
            level = 95,
            alpha = 0.5)
```


```{r entr-cv, include=FALSE}
training <- y %>% filter(month(index) <= 11)
test <- y %>% filter(month(index) > 11)

train_fit <- training %>%
  model(
    sarimacv = ARIMA(box_cox(Ozone, lambda) ~ 0 + pdq(1, 0, 1) + PDQ(0, 1, 2, period=24))
  )

test_refit <- train_fit %>% refit(test)
# fits5 <- test_refit %>% fitted(h = 5)
# saveRDS(fits5, file = "models/entrecampos_cv_fits5.rds")
# Load
fits5 <- readRDS("models/entrecampos_cv_fits5.rds")

#fits1 <- test_refit %>% fitted(h = 1)
```

```{r entr-cv-accuracy-tab}
kbl(bind_rows(
      fit %>% accuracy() %>% select(RMSE, MAE),
      test_refit %>% accuracy() %>% select(RMSE, MAE)
      ) %>% 
      add_column(Type = c("Series", "CV")) %>% 
      select(3, 1, 2),
    caption = paste(y$Location[1], format(test_refit %>% pull(2)), "Series and Crossvalidation Accuracy"),
    booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```


```{r entr-cv-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(test_refit %>% pull(2)), "crossvalidation forecasts"), fig.dim=c(6, 2), include=TRUE}
test %>%
  autoplot(Ozone) +
  autolayer(fits5, .fitted, col = "#D55E00")
```

