---
title: "Estarreja"
author: "helder"
date: "`r Sys.Date()`"
output: pdf_document
---

## Estarreja

```{r , include=FALSE}
y <- qualityAR03 %>% filter(Location=="Estarreja")
```

```{r, echo=FALSE, include=FALSE}
lambda <- y %>%
  features(Ozone, features = guerrero) %>%
  pull(lambda_guerrero)
#caption_tex = paste0("Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2))
```

The Estarreja series is ploted in figure \@ref(fig:est-ts). The ACF an PACF plots show non-stationarity, and a seasonal component at lags 24, 48, etc. An Order 1 Difference in seasonal component of SARIMA model will be used and a BOX COX tranformation to improve stability.

Again, applying Unit root tests as resumed in table \ref{tab:est-test-stationary} leads to incorrect assumptions, of for what type of difference should by applied.

```{r est-ts, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Time series"), fig.dim=c(7, 3), include=TRUE}
y %>% gg_tsdisplay(Ozone, plot_type = "partial", lag_max = 200)
```

```{r est-test-stationary, include=TRUE}
adf <- function(x, ...) {
  out <- tseries::adf.test(x, k = 1)
  c(adf_stat = unname(out$statistic), adf_pvalue = out$p.value)
}
# tseries::adf.test(box_cox(y$Ozone, lambda), k=1)
kbl(y %>%
      mutate(Ozone = box_cox(Ozone, lambda)) %>%
      features(Ozone, 
               list(unitroot_kpss, adf, unitroot_ndiffs, ~ unitroot_nsdiffs(., .period = 24))),
    caption = paste(y$Location[1], " - Unit root tests and suggested number of differences"),
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Looking at the seasonal difference series (h=24) in figure \@ref(fig:est-diffplot), it shows no trend, and the ACF and PACF plots look typical of a stationary series (no further differentiate). There are two seasonal significant correlation at lag 24 and 48 in the ACF, meaning that an SMA(2) component model will be used. In the the first lags only 2 are significant in the PACF and a AR(2) component will be considered. The final predicted model is of the form SARIMA(2,0,0)(0,1,2)[24]. A wide range of variations around this form were considered, to check if there was any improvement in goodness-of-fit measures, as resumed in  table \@ref(tab:est-model-comparison).

```{r est-diffplot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2)), fig.dim=c(7, 4), include=TRUE}
y %>%
  gg_tsdisplay(difference(box_cox(Ozone, lambda), lag = 24), plot_type = "partial", lag_max = 200) +
  labs(y = "")
```

```{r , include=FALSE}
#  [1] "Location"     "sarima000101" "sarima100101" "sarima001101" "sarima101101" "sarima201101"
#  [7] "sarima102101" "sarima202101" "sarima000012" "sarima100012" "sarima200012" "sarima101012"
# [13] "sarima201012"
#fits <- readRDS("models/estarreja_fit_s012.rds") %>% select(-c(2, 3, 4, 5, 6, 7, 8, 9))
fits <- readRDS("models/estarreja_fit_s012.rds")
fit <- fits %>% select(Location, sarima200012)
```

Diagnostics in figure \@ref(fig:est-fit-resid-plot) show that predicted model pass the test for non-correlated residuals. But again the QQ plot shows patterns diverging from normality in the tails of the distribution. The Fitted vs Residual plot does not show any pattern of concern, as with the histogram.

The alternative fitted on other locations, <ARIMA(1,0,1)(0,1,2)[24]>, have correlated residuals, so in this location it will not be used. The decision is to keep the predicted model `r format(fit %>% pull(2))`, only raising concern by not following a normal distribution as seen in the qq plot. Other variations do not show improvement in model AIC, AICc or BIC.

```{r, include=FALSE}
plot.box.ljung <- function (
    z, k = 10, main = "Ljung-Box test (k)", ylab = "p-value"
) {
  p <- rep(NA, k)

  for (i in 1:k) {
    p[i] <- Box.test(z, i, type = "Ljung-Box")$p.value
  }

  ggplot(as_tibble(p), aes(x=1:k, y=p)) +
    geom_bar(stat='identity', width =0.05) +
    geom_point() + ylim(0,1) + ylab(ylab) +
    geom_abline(intercept = .05, slope = 0, lty = 3) +
    ggtitle(main)
}

fit_diag_plots <- function(fit) {
  m <- fit %>% augment() %>% mutate( .innov = scale(.innov) )

  p1 <- m %>% autoplot(.innov)
  p2 <- m %>% ACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p3 <- m %>% PACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p4 <- m %>%  gghistogram(x=".innov", bins = 150)
  p5 <- plot.box.ljung(m %>% pull(.innov), k=10)
  p6 <- ggqqplot(m,".innov", title = "QQ plot")
  p7 <- ggplot(m, aes(x = .fitted, y = .innov)) +
          geom_point()
  
  (p1) / (p2 + p3) / (p4 + p5) / (p6 + p7)
}

```

```{r est-fit-resid-plot, fig.pos="H", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE}
fit_diag_plots(fit)
```

```{r est-model-comparison, include=TRUE}
kbl(fits %>% 
      pivot_longer(!Location, names_to = "Model name",
                     values_to = "Orders") %>%
      mutate("Model" = format(Orders)) %>%
      glance() %>%
      select(Location, Model, AIC, AICc, BIC),
    format = "latex", 
    caption = paste(y$Location[1], "model comparison"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

The coefficients are statistical significant, as resumed in table \@ref(tab:est-coef), all p-values of the t-test (H0: coefficient is zero) lead to rejecting the null hypothesis. In this location the process also evolves with positive correlation with the last value and negative correlation with last seasonal period, but have a small negative correlation with lag 2. In the previous locations the sign of the correlations were consistent here the correlation with lag 2 is in opposite sign than the lag 1.

```{r est-coef, include=TRUE}
kbl(fit %>% coef(), 
    format = "latex", 
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Coefficient statistics"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Model forecasts were obtained in the table \@ref(tab:est-forecast-tab), and figure \@ref(fig:est-forecast-plot). This model have smaller variance in the innovation process than previous ones.

Cross-validation on the last month shows that the model is capturing the time dependency structure of the series. The dynamic range is maintained even some sustained low values are foretasted (normal innovations do not produce sustained high/low values), but sustained high values are not forecasted in the the same pattern.

```{r est-forecast-tab, include=TRUE}
f <- fit %>% forecast(h=5)

kbl(f %>%  hilo(level = c(95)) %>% select(-c(1,2)),
             caption = paste(y$Location[1], format(fit %>% pull(2)), "forecast 5 periods ahead (95\\% CI)"),
             booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r est-forecast-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "forecasts"), fig.dim=c(6, 2), include=TRUE}
f %>% 
  autoplot(y %>%
              filter(between(index,as_datetime("2020-12-15"), as_datetime("2021-01-01"))),
            level = 95,
            alpha = 0.5)
```


```{r est-cv, include=FALSE}
training <- y %>% filter(month(index) <= 11)
test <- y %>% filter(month(index) > 11)

train_fit <- training %>%
  model(
    sarimacv = ARIMA(box_cox(Ozone, lambda) ~ 0 + pdq(2, 0, 0) + PDQ(0, 1, 2, period=24))
  )

test_refit <- train_fit %>% refit(test)
# fits5 <- test_refit %>% fitted(h = 5)
# saveRDS(fits5, file = "models/estarreja_cv_fits5.rds")
# Load
fits5 <- readRDS("models/estarreja_cv_fits5.rds")
```

```{r est-cv-accuracy-tab}
kbl(bind_rows(
      fit %>% accuracy() %>% select(RMSE, MAE),
      test_refit %>% accuracy() %>% select(RMSE, MAE)
      ) %>% 
      add_column(Type = c("Series", "CV")) %>% 
      select(3, 1, 2),
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Series and Crossvalidation Accuracy"),
    booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r est-cv-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(test_refit %>% pull(2)), "crossvalidation forecasts"), fig.dim=c(6, 2), include=TRUE}
test %>%
  autoplot(Ozone) +
  autolayer(fits5, .fitted, col = "#D55E00")
```

