---
title: "Ilhavo"
author: "helder"
date: "`r Sys.Date()`"
output: pdf_document
---

## Ilhavo {-}

```{r , include=FALSE}
# Ihavo (SIC)
y <- qualityAR03 %>% filter(Location=="Ihavo")
```

```{r, echo=FALSE, include=FALSE}
lambda <- y %>%
  features(Ozone, features = guerrero) %>%
  pull(lambda_guerrero)
#caption_tex = paste0("Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2))
```

The Ilhavo series is ploted in figure \@ref(fig:ilh-ts). The ACF an PACF plots show non-stationarity, and a seasonal component at lags 24, 48, etc. There are some fluctuations in the variance and a Box Cox transformation will be applied. To remove seasonality an order 1 Difference will be used. 

Unit root tests as resumed in table \ref{tab:ilh-esth-stationary} suggesting additional differencing that is not needed.

```{r ilh-ts, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Time series"), fig.dim=c(7, 3), include=TRUE}
y %>% gg_tsdisplay(Ozone, plot_type = "partial", lag_max = 200)
```

```{r ilh-esth-stationary, include=TRUE}
adf <- function(x, ...) {
  out <- tseries::adf.test(x, k = 1)
  c(adf_stat = unname(out$statistic), adf_pvalue = out$p.value)
}
# tseries::adf.test(box_cox(y$Ozone, lambda), k=1)
kbl(y %>%
      mutate(Ozone = box_cox(Ozone, lambda)) %>%
      features(Ozone, 
               list(unitroot_kpss, adf, unitroot_ndiffs, ~ unitroot_nsdiffs(., .period = 24))),
    caption = paste(y$Location[1], " - Unit root tests and suggested number of differences"),
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Looking at the seasonal difference series (h=24) in figure \@ref(fig:ilh-diffplot), it shows no trend, and the ACF and PACF plots look typical of a stationary series (no further differentiate). There are two seasonal significant correlation at lag 24 and 48, in the ACF meaning that an SMA(2) component should considered. The first lags of the PACF have only 2 that are significant implying an AR(2) component. The predicted model is of the form SARIMA(2,0,0)(0,1,2)[24].

```{r ilh-diffplot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2)), fig.dim=c(7, 4), include=TRUE}
y %>%
  gg_tsdisplay(difference(box_cox(Ozone, lambda), lag = 24), plot_type = "partial", lag_max = 200) +
  labs(y = "")
```

```{r load-fit, include=FALSE}
#  [1] "Location"     "sarima000101" "sarima100101" "sarima001101" "sarima101101" "sarima201101"
#  [7] "sarima102101" "sarima202101" "sarima000012" "sarima100012" "sarima200012" "sarima101012"
# [13] "sarima201012"
#fits <- readRDS("models/Ilhavo_fit_s012.rds") %>% select(-c(2, 3, 4, 5, 6, 7, 8, 9))
fits <- readRDS("models/ilhavo_fit_s012.rds")
fit <- fits %>% select(Location, sarima202012)
```

Analysis the predicted model, lead to conclude the residuals are correlated, leading to search variations. Model goodness of fit comparison is resumed in table \@ref(tab:ilh-model-comparison). From candidates, model <ARIMA(2,0,1)(0,1,2)[24]> also resulted in correlated residuals, and only <ARIMA(2,0,2)(0,1,2)[24]> had uncorrelated residuals.

Diagnostics plots of this model are presented in figure \@ref(fig:ilh-fit-resid-plot). Apart from statistical significance of nocorrelated residuals in higher lags, the QQ plot shows patterns diverging from normality in the tails of the distribution. The Fitted vs Residual and histogram look good.

```{r, include=FALSE}
plot.box.ljung <- function (
    z, k = 10, main = "Ljung-Box test (k)", ylab = "p-value"
) {
  p <- rep(NA, k)

  for (i in 1:k) {
    p[i] <- Box.test(z, i, type = "Ljung-Box")$p.value
  }

  ggplot(as_tibble(p), aes(x=1:k, y=p)) +
    geom_bar(stat='identity', width =0.05) +
    geom_point() + ylim(0,1) + ylab(ylab) +
    geom_abline(intercept = .05, slope = 0, lty = 3) +
    ggtitle(main)
}
```

```{r ilh-fit-resid-plot, fig.pos="H", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE}
m <- fit %>% augment() %>% mutate( .innov = scale(.innov) )

p1 <- m %>% autoplot(.innov)
p2 <- m %>% ACF(.innov, lag_max = 60) %>% autoplot() + ylim(-0.1, 0.3)
p3 <- m %>% PACF(.innov, lag_max = 60) %>% autoplot() + ylim(-0.1, 0.3)
p4 <- m %>%  gghistogram(x=".innov", bins = 150)
p5 <- plot.box.ljung(m %>% pull(.innov), k=10)
p6 <- ggqqplot(m,".innov", title = "QQ plot")
p7 <- ggplot(m, aes(x = .fitted, y = .innov)) +
        geom_point()

(p1) / (p2 + p3) / (p4 + p5) / (p6 + p7)
rm(m)
```

```{r ilh-model-comparison, include=TRUE}
kbl(fits %>% 
      pivot_longer(!Location, names_to = "Model name",
                     values_to = "Orders") %>%
      mutate("Model" = format(Orders)) %>%
      glance() %>%
      select(Location, Model, AIC, AICc, BIC),
    format = "latex", 
    caption = paste(y$Location[1], "model comparison"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

```{r, include=FALSE, eval=FALSE}
fit <- fits %>% select(Location, sarima201012)
```

```{r ilh-fit-resid-plot0, fig.pos="H", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE, eval=FALSE}
m <- fit %>% augment() %>% mutate( .innov = scale(.innov) )

p1 <- m %>% autoplot(.innov)
p2 <- m %>% ACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
p3 <- m %>% PACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
p4 <- m %>%  gghistogram(x=".innov", bins = 150)
p5 <- plot.box.ljung(m %>% pull(.innov), k=10)
p6 <- ggqqplot(m,".innov", title = "QQ plot")
p7 <- ggplot(m, aes(x = .fitted, y = .innov)) +
        geom_point()

(p1) / (p2 + p3) / (p4 + p5) / (p6 + p7)
rm(m)
```

The coefficients are statistical significant, as resumed in table \@ref(tab:ilh-coef), all p-values of the t-test (H0: coefficient is zero) lead to rejecting the null hypothesis. The process is drove by positive correlations with last value by negative correlation to the moving average component and negative correlations with the last seasonal period.

```{r ilh-coef, include=TRUE}
kbl(fit %>% coef(), 
    format = "latex", 
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Coefficient statistics"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Model forecasts were obtained in the table \@ref(tab:ilh-forecast-tab), and figure \@ref(fig:ilh-forecast-plot). The sample path of the process in recent lags is not realistic and the forecast is not meaningful. This model also have smaller variance in the innovation process than the first two.

Cross-validation on the last month shows that the model lacks capturing the time dependency structure of the series following sustained high values, this may due the prediction horizon of h=5.

```{r ilh-forecast-tab, include=TRUE}
f <- fit %>% forecast(h=5)

kbl(f %>%  hilo(level = c(95)) %>% select(-c(1,2)),
             caption = paste(y$Location[1], format(fit %>% pull(2)), "forecast 5 periods ahead (95\\% CI)"),
             booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r ilh-forecast-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "forecasts"), fig.dim=c(6, 2), include=TRUE}
f %>% 
  autoplot(y %>%
              filter(between(index,as_datetime("2020-12-15"), as_datetime("2021-01-01"))),
            level = 95,
            alpha = 0.5)
```

```{r ilh-cv, include=FALSE}
training <- y %>% filter(month(index) <= 11)
test <- y %>% filter(month(index) > 11)

train_fit <- training %>%
  model(
    sarimacv = ARIMA(box_cox(Ozone, lambda) ~ 0 + pdq(2, 0, 2) + PDQ(0, 1, 2, period=24))
  )

test_refit <- train_fit %>% refit(test)

#fits5 <- test_refit %>% fitted(h = 5)
#saveRDS(fits5, file = "models/ilhavo_cv_fits5.rds")
# Load the city object as city
fits5 <- readRDS("models/ilhavo_cv_fits5.rds")

#fits1 <- test_refit %>% fitted(h = 1)
```

```{r ilh-cv-accuracy-tab}
kbl(bind_rows(
      fit %>% accuracy() %>% select(RMSE, MAE),
      test_refit %>% accuracy() %>% select(RMSE, MAE)
      ) %>% 
      add_column(Type = c("Series", "CV")) %>% 
      select(3, 1, 2),
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Series and Crossvalidation Accuracy"),
    booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r ilh-cv-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(test_refit %>% pull(2)), "crossvalidation forecasts"), fig.dim=c(6, 2), include=TRUE}
test %>%
  autoplot(Ozone) +
  autolayer(fits5, .fitted, col = "#D55E00")
```

