---
title: "Restelo"
author: "helder"
date: "`r Sys.Date()`"
output: pdf_document
---

## Restelo {-}

```{r , include=FALSE}
# Ihavo (SIC)
y <- qualityAR03 %>% filter(Location=="Restelo")
```

```{r, echo=FALSE, include=FALSE}
lambda <- y %>%
  features(Ozone, features = guerrero) %>%
  pull(lambda_guerrero)
#caption_tex = paste0("Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2))
```

The Restelo series plots are presented in figure \@ref(fig:res-ts). A Box-Cox transformation was applied and also a seasonal difference is needed.

The suggestions of differencing of unit root tests as resumed in table \ref{tab:res-esth-stationary} were discarded.

```{r res-ts, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Time series"), fig.dim=c(7, 3), include=TRUE}
y %>% gg_tsdisplay(Ozone, plot_type = "partial", lag_max = 200)
```

```{r res-esth-stationary, include=TRUE}
adf <- function(x, ...) {
  out <- tseries::adf.test(x, k = 1)
  c(adf_stat = unname(out$statistic), adf_pvalue = out$p.value)
}
# tseries::adf.test(box_cox(y$Ozone, lambda), k=1)
kbl(y %>%
      mutate(Ozone = box_cox(Ozone, lambda)) %>%
      features(Ozone, 
               list(unitroot_kpss, adf, unitroot_ndiffs, ~ unitroot_nsdiffs(., .period = 24))),
    caption = paste(y$Location[1], " - Unit root tests and suggested number of differences"),
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

The seasonal difference series (h=24) in figure \@ref(fig:res-diffplot), shows only one seasonal significant correlation at lag 24, maybe also at lag 48 (in the ACF) and only two first lags significant (in the PACF). The predicted model is of the form SARIMA(2,0,0)(0,1,1)[24].

```{r res-diffplot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2)), fig.dim=c(7, 4), include=TRUE}
y %>%
  gg_tsdisplay(difference(box_cox(Ozone, lambda), lag = 24), plot_type = "partial", lag_max = 200) +
  labs(y = "")
```

```{r , include=FALSE}
#  [1] "Location"     "sarima000101" "sarima100101" "sarima001101" "sarima101101" "sarima201101"
#  [7] "sarima102101" "sarima202101" "sarima000012" "sarima100012" "sarima200012" "sarima101012"
# [13] "sarima201012"
#fits <- readRDS("models/Laranjeiro-Alamada_fit_s012.rds") %>% select(-c(2, 3, 4, 5, 6, 7, 8, 9))
fits <- readRDS("models/restelo_fit_s011.rds")
fit <- fits %>% select(Location, sarima301011) # is 202012, there is a naming error in save!
```

Variations where considered and resumed in table \@ref(tab:res-model-comparison). Some of the models pass the non-correlated residuals test: sarima301011, sarima202012 and sarima301012.

The model `r format(fit %>% pull(2))` was selected because of the goodness-of-fit measures and also was selected to some of the other locations. The diagnostics plots are in figure \@ref(fig:res-fit-resid-plot). The most concerning is the the QQ plot showing patterns diverging from normality in the tails of the distribution.

```{r, include=FALSE}
plot.box.ljung <- function (
    z, k = 10, main = "Ljung-Box test (k)", ylab = "p-value"
) {
  p <- rep(NA, k)

  for (i in 1:k) {
    p[i] <- Box.test(z, i, type = "Ljung-Box")$p.value
  }

  ggplot(as_tibble(p), aes(x=1:k, y=p)) +
    geom_bar(stat='identity', width =0.05) +
    geom_point() + ylim(0,1) + ylab(ylab) +
    geom_abline(intercept = .05, slope = 0, lty = 3) +
    ggtitle(main)
}

fit_diag_plots <- function(fit) {
  m <- fit %>% augment() %>% mutate( .innov = scale(.innov) )

  p1 <- m %>% autoplot(.innov)
  p2 <- m %>% ACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p3 <- m %>% PACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p4 <- m %>%  gghistogram(x=".innov", bins = 150)
  p5 <- plot.box.ljung(m %>% pull(.innov), k=10)
  p6 <- ggqqplot(m,".innov", title = "QQ plot")
  p7 <- ggplot(m, aes(x = .fitted, y = .innov)) +
          geom_point()
  
  (p1) / (p2 + p3) / (p4 + p5) / (p6 + p7)
}

```


```{r res-fit-resid-plot, fig.pos="H", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE}
fit_diag_plots(fit)
```

```{r res-model-comparison, include=TRUE}
kbl(fits %>% 
      pivot_longer(!Location, names_to = "Model name",
                     values_to = "Orders") %>%
      mutate("Model" = format(Orders)) %>%
      glance() %>%
      select(Location, Model, AIC, AICc, BIC),
    format = "latex", 
    caption = paste(y$Location[1], "model comparison"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```


```{r, include=FALSE, eval=FALSE}
fit <- fits %>% select(Location, sarima201012)
```

```{r res-fit-resid-plot0, fig.pos="H", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE, eval=FALSE}
fit_diag_plots(fit)
```

The coefficients are statistical significant, as resumed in table \@ref(tab:res-coef), all p-values of the t-test (H0: coefficient is zero) lead to rejecting the null hypothesis.

```{r res-coef, include=TRUE}
kbl(fit %>% coef(), 
    format = "latex", 
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Coefficient statistics"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Model forecasts were obtained in the table \@ref(tab:res-forecast-tab), and figure \@ref(fig:res-forecast-plot). Cross-validation on the last month shows that the model capturing the time dependency structure of the series. The dynamic range is maintained even some sustained high/low values are forecasted (normal innovations do not produce sustained high/low values).

```{r res-forecast-tab, include=TRUE}
f <- fit %>% forecast(h=5)

kbl(f %>%  hilo(level = c(95)) %>% select(-c(1,2)),
             caption = paste(y$Location[1], format(fit %>% pull(2)), "forecast 5 periods ahead (95\\% CI)"),
             booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r res-forecast-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "forecasts"), fig.dim=c(6, 2), include=TRUE}
f %>% 
  autoplot(y %>%
              filter(between(index,as_datetime("2020-12-15"), as_datetime("2021-01-01"))),
            level = 95,
            alpha = 0.5)
```

```{r res-cv, include=FALSE}
# training <- y %>% filter(month(index) <= 11)
# test <- y %>% filter(month(index) > 11)
# 
# train_fit <- training %>%
#   model(
#     sarimacv = ARIMA(box_cox(Ozone, lambda) ~ 0 + pdq(3, 0, 1) + PDQ(0, 1, 1, period=24))
#   )
# 
# test_refit <- train_fit %>% refit(test)

#fits5 <- test_refit %>% fitted(h = 5)
#saveRDS(fits5, file = "models/restelo_cv_fits5.rds")
# saveRDS(test_refit, file = "models/restelo_cv_test_refit.rds")
# Load
fits5 <- readRDS("models/restelo_cv_fits5.rds")
test_refit <- readRDS("models/restelo_cv_test_refit.rds")
#fits1 <- test_refit %>% fitted(h = 1)
```

```{r res-cv-accuracy-tab}
kbl(bind_rows(
      fit %>% accuracy() %>% select(RMSE, MAE),
      test_refit %>% accuracy() %>% select(RMSE, MAE)
      ) %>% 
      add_column(Type = c("Series", "CV")) %>% 
      select(3, 1, 2),
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Series and Crossvalidation Accuracy"),
    booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r res-cv-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(test_refit %>% pull(2)), "crossvalidation forecasts"), fig.dim=c(6, 2), include=TRUE}
test %>%
  autoplot(Ozone) +
  autolayer(fits5, .fitted, col = "#D55E00")
```

