---
title: "VNTelha-Maia"
author: "helder"
date: "`r Sys.Date()`"
output: pdf_document
---

## VNTelha-Maia {-}

```{r , include=FALSE}
y <- qualityAR03 %>% filter(Location=="VNTelha-Maia")
```

```{r, echo=FALSE, include=FALSE}
lambda <- y %>%
  features(Ozone, features = guerrero) %>%
  pull(lambda_guerrero)
#caption_tex = paste0("Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2))
```

The VNTelha-Maia series looks like figure \@ref(fig:vnt-ts). There are fluctuations in the variance so a Box Cox transformation will be applied. The ACF an PACF plots show non-stationarity, and a seasonal component at lags 24, 48, etc. An Order 1 Difference in seasonal component of SARIMA model will be used. 

Unit root tests as resumed in table \ref{tab:vnt-esth-stationary} sugestion of differencing the series in its own frequency will not be followed.

```{r vnt-ts, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Time series"), fig.dim=c(7, 3), include=TRUE}
y %>% gg_tsdisplay(Ozone, plot_type = "partial", lag_max = 200)
```

```{r vnt-esth-stationary, include=TRUE}
adf <- function(x, ...) {
  out <- tseries::adf.test(x, k = 1)
  c(adf_stat = unname(out$statistic), adf_pvalue = out$p.value)
}
# tseries::adf.test(box_cox(y$Ozone, lambda), k=1)
kbl(y %>%
      mutate(Ozone = box_cox(Ozone, lambda)) %>%
      features(Ozone, 
               list(unitroot_kpss, adf, unitroot_ndiffs, ~ unitroot_nsdiffs(., .period = 24))),
    caption = paste(y$Location[1], " - Unit root tests and suggested number of differences"),
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

The seasonal difference series (h=24) in figure \@ref(fig:vnt-diffplot), analysis of ACF shows significant correlation at lag 24, SMA(1) component. In the PACF the first lags only 2 are significant, so AR(2) component will be considered. The final predicted model is of the form SARIMA(2,0,0)(0,1,2)[24].

```{r vnt-diffplot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], " - Difference h=24 Box-Cox Transformation $\\lambda$ = ", round(lambda,2)), fig.dim=c(7, 4), include=TRUE}
y %>%
  gg_tsdisplay(difference(box_cox(Ozone, lambda), lag = 24), plot_type = "partial", lag_max = 200) +
  labs(y = "")
```

```{r , include=FALSE}
fits <- readRDS("models/vntelha-maia_fit_s012.rds")
fit <- fits %>% select(Location, sarima101011)
```


After testing variations of the predicted model, resumed in table \@ref(tab:vnt-model-comparison), the best model is the `r #format(fit %>% pull(2))`. One of the simplest forms.

Models diagnostics plots in figure \@ref(fig:vnt-fit-resid-plot), confirm non-correlated residuals.

```{r, include=FALSE}
plot.box.ljung <- function (
    z, k = 10, main = "Ljung-Box test (k)", ylab = "p-value"
) {
  p <- rep(NA, k)

  for (i in 1:k) {
    p[i] <- Box.test(z, i, type = "Ljung-Box")$p.value
  }

  ggplot(as_tibble(p), aes(x=1:k, y=p)) +
    geom_bar(stat='identity', width =0.05) +
    geom_point() + ylim(0,1) + ylab(ylab) +
    geom_abline(intercept = .05, slope = 0, lty = 3) +
    ggtitle(main)
}

fit_diag_plots <- function(fit) {
  m <- fit %>% augment() %>% mutate( .innov = scale(.innov) )

  p1 <- m %>% autoplot(.innov)
  p2 <- m %>% ACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p3 <- m %>% PACF(.innov, lag_max = 100) %>% autoplot() + ylim(-0.1, 0.2)
  p4 <- m %>%  gghistogram(x=".innov", bins = 150)
  p5 <- plot.box.ljung(m %>% pull(.innov), k=10)
  p6 <- ggqqplot(m,".innov", title = "QQ plot")
  p7 <- ggplot(m, aes(x = .fitted, y = .innov)) +
          geom_point()
  
  (p1) / (p2 + p3) / (p4 + p5) / (p6 + p7)
}

```


```{r vnt-fit-resid-plot, fig.pos="H", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "standardized residuals diagnostic plots"), fig.dim=c(7, 8), include=TRUE}
fit_diag_plots(fit)
```

```{r vnt-model-comparison, include=TRUE}
kbl(fits %>% 
      pivot_longer(!Location, names_to = "Model name",
                     values_to = "Orders") %>%
      mutate("Model" = format(Orders)) %>%
      glance() %>%
      select(Location, Model, AIC, AICc, BIC),
    format = "latex", 
    caption = paste(y$Location[1], "model comparison"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

The coefficients are statistical significant, as resumed in table \@ref(tab:vnt-coef), all p-values of the t-test (H0: coefficient is zero) lead to rejecting the null hypothesis.

```{r vnt-coef, include=TRUE}
kbl(fit %>% coef(), 
    format = "latex", 
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Coefficient statistics"), 
    booktabs = T) %>% 
  kable_styling(latex_options = "hold_position")
```

Model forecasts were obtained in the table \@ref(tab:vnt-forecast-tab), and figure \@ref(fig:vnt-forecast-plot). Cross-validation (horizon = 5) on the last month shows that the dynamic range not followed by the model missing lower values and producing higher peaks.

```{r vnt-forecast-tab, include=TRUE}
f <- fit %>% forecast(h=5)

kbl(f %>%  hilo(level = c(95)) %>% select(-c(1,2)),
             caption = paste(y$Location[1], format(fit %>% pull(2)), "forecast 5 periods ahead (95\\% CI)"),
             booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r vnt-forecast-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(fit %>% pull(2)), "forecasts"), fig.dim=c(6, 2), include=TRUE}
f %>% 
  autoplot(y %>%
              filter(between(index,as_datetime("2020-12-15"), as_datetime("2021-01-01"))),
            level = 95,
            alpha = 0.5)
```

```{r vnt-cv, include=FALSE}
training <- y %>% filter(month(index) <= 11)
test <- y %>% filter(month(index) > 11)

train_fit <- training %>%
  model(
    sarimacv = ARIMA(box_cox(Ozone, lambda) ~ 0 + pdq(1, 0, 1) + PDQ(0, 1, 1, period=24))
  )

test_refit <- train_fit %>% refit(test)

# fits5 <- test_refit %>% fitted(h = 5)
# saveRDS(fits5, file = "models/VNTelha-Maia_cv_fits5.rds")
# Load
fits5 <- readRDS("models/VNTelha-Maia_cv_fits5.rds")

#fits1 <- test_refit %>% fitted(h = 1)
```

```{r vnt-cv-accuracy-tab}
kbl(bind_rows(
      fit %>% accuracy() %>% select(RMSE, MAE),
      test_refit %>% accuracy() %>% select(RMSE, MAE)
      ) %>% 
      add_column(Type = c("Series", "CV")) %>% 
      select(3, 1, 2),
    caption = paste(y$Location[1], format(fit %>% pull(2)), "Series and Crossvalidation Accuracy"),
    booktabs = T) %>%
  kable_styling(latex_options = "hold_position")
```

```{r vnt-cv-plot, fig.pos="!h", out.extra = "", fig.cap=paste(y$Location[1], format(test_refit %>% pull(2)), "crossvalidation forecasts"), fig.dim=c(6, 2), include=TRUE}
test %>%
  autoplot(Ozone) +
  autolayer(fits5, .fitted, col = "#D55E00")
```

